### 2014-10-31 implementation.tex
Minor edits to implementation section

Replicated data type implementation and the stronger consistency semantics are
enforced in the \emph{shim layer}. Our implementation supports eventual, causal
and strong consistency for operations, and RC, MAV and RR transaction
semantics. All of this functionality is implemented on top of the standard
interface exposed by Cassandra, and illustrates the portability of \name
programming model. Complex business logic is implemented with the help of
transactions on the server side, with clients interacting with the web-service
over REST interface. From an engineering perspective, resorting to an
off-the-shelf backing store meant that our entire implementation comprised of
only around 2500 lines of Haskell code, which is packaged as a Haskell library.

@@

Every effect generated as a result of an effectful operation on an object
inserts a new row $(o,e,vis,txn,val)$ into the backing store, where $o$ and $e$
are object and (unique) effect identifiers, $vis$ is the set of identifiers of
effects visible to this operation, $txn$ is an optional transaction identifier,
and $val$ is the value associated with the effect. The shim layer periodically
fetches updates from the backing store for those objects which were accessed in
the previous time slice. Since causally consistent operations require an
up-to-date view of the current session, the shim layer node synchronously
fetches operations if the causally preceding operations in the current session
are not available in the cache. Strongly consistent operations are performed
after obtaining exclusive leases on objects. The lease mechanism is implemented
with the help of Cassandra's support for conditional updates and expiring
columns.

@@

For every transaction, we instantiate a special transaction marker effect $m$,
but importantly, do not insert into the backing store. $m$ is included as a
dependence to every effect (let $e$ be one such effect) generated in the
transaction. Since $m$ has not yet been written to the store, no operation will
witness $e$ while the transaction in progress. Let $S$ be the set of effects
generated in the transaction. At the end of the transaction, we insert $m$ into
the backing store with $S$ as its dependence. Now any replica which includes
$e$, must include $m$, and transitively must include every effect from the
transaction. This ensures atomicity and satisfies RC requirement.

>>>>

Replicated data type implementation and the stronger consistency semantics are
enforced in the \emph{shim layer}. Our implementation supports eventual, causal
and strong consistency for operations, and RC, MAV and RR transaction
semantics. All of this functionality is implemented on top of the standard
interface exposed by Cassandra, and illustrates the portability of \name
programming model. Complex business logic is implemented with the help of
transactions on the web-server side, with clients interacting with the
web-service over REST interface. From an engineering perspective, resorting to
an off-the-shelf backing store meant that our entire implementation comprised of
only around 2500 lines of Haskell code, which is packaged as a Haskell
library\footnote{url to the hackage}.

@@

Every effect generated as a result of an effectful operation on an object
inserts a new row $(o,e,vis,txn,val)$ into the backing store, where $o$ and $e$
are object and (unique) effect identifiers, $vis$ is the set of identifiers of
effects visible to this operation, $txn$ is an optional transaction identifier,
and $val$ is the value associated with the effect (eg: \cf{Withdraw 50}). The
shim layer periodically fetches updates from the backing store for those objects
which were accessed since updates were last fetched. Since causally consistent
operations require an up-to-date view of the current session, the shim layer
node synchronously fetches operations if the causally preceding operations in
the current session are not available in the cache.  Strongly consistent
operations are performed after obtaining exclusive leases on objects. The lease
mechanism is implemented with the help of Cassandra's support for conditional
updates and expiring columns.

@@

For every transaction, we instantiate a special transaction marker effect $m$,
but importantly, do not insert into the backing store. $m$ is included as a
dependence to every effect (let $e$ be one such effect) generated in the
transaction. Since $m$ has not yet been written to the store, no operation will
witness $e$ while the transaction in progress. Let $S$ be the set of effects
generated in the transaction. At the end of the transaction, we insert $m$ into
the backing store with $S$ as its dependence. Now any replica which includes
$e$, must include $m$, and transitively must include every effect from the
transaction. This ensures atomicity and satisfies the RC requirement.
